True versus apparent shapes of astrophysical bow shocks

### 2016-05-06 papers/Paper1/sec-conic.tex
Improve R90 description. Unfinished
Corrections to allow paper to compile
Tidy up the conic section of paper
Remove an unwanted figure and rearrange the others.  Remove extraneous
blank lines in latex file

\section{Conic section approximation to bow shock shapes}
\label{sec:conic}

@@

%Instead of the excentricity, we utilize the parameter $\theta_c$ to characterize the different curves, where
%$\tan\theta_c = \frac{b}{a}$,  $b$ and $a$ are the typical parameters of conics. A positive value for $\theta_c$ indicates that the given curve is a closed one, i.e
%an ellipse, while a negative value indicates that is an hyperbola. %Insert figures if neccesary  
\begin{figure}
\begin{tabular}{cc}
\includegraphics[width=0.4\linewidth]{ellipse} &
\includegraphics[width=0.4\linewidth]{hiperbola}
\end{tabular}
\label{fig:conics}
\caption{Schematic representation of conic like bowshocks: A) Ellipse and B) Hyperbola The center of the conic is located at a distance $x_0$ from the origin at the $x$ axis. The shell position is 
described either by cartesian coordinates $(x,y)$ or polar coordinates $(R,\theta)$. In this section is more convenient to use cartesian coordinates, as their parametrization is quite easy.}
\end{figure}
Due to the similarities between the parametrization of the ellipse and the hyperbola, we can do the following parametrization:

\begin{align}
x = a\Cos(t)-x_0 \\ 
y = b\Sin(t)
\end{align}

where:
\begin{align}
\Cos(t) = \left\lbrace \begin{array}{c}
\cos t ~\mathrm{if~ellipse} \\
\cosh t ~\mathrm{if~hyperbola}
\end{array}\right.\\
\Sin(t) = \left\lbrace \begin{array}{c}
\sin t ~\mathrm{if~ellipse} \\
\sinh t ~\mathrm{if~hyperbola}
\end{array}\right. \\
-\pi < t < \pi \\
R_0 = a - x_0 \\
a<0 ~ \mathrm{for~hyperbola}
\end{align}
We have set a whole family of curves with a singular parameter denoted as $\tan\theta_c \equiv \frac{b}{a}$. Positive values of this parameter describe closed curves
(i.e. ellipses) and the negative ones describe open curves (i.e. hyperbolas). Particular cases are $\theta_c =\frac{\pi}{4}$, which describes a circle and $\theta_c=0$, which describes
a parabola.
\begin{figure}
\includegraphics[width=\linewidth]{conic1}
\caption{A whole set of conics with the same radius of curvature. The parameter $\theta_c$ is used as the equivalent of the excentricity. This angle is defined as positive for closed curves
and negative for the opened ones. An special case is $\theta_c=0$, which describes a parabola.}
\label{fig:conics-family}
\end{figure}
\subsection{Projection onto the plane of sky}

Once the parametrization is done, we can find the apparent shape of the shell in the observer's frame, following the procedure explained in section \ref{sec:projection}.

First of all, the intrinsic 3D shape of the shell is given by:

\begin{align}
x = a\Cos(t)-x_0 \\ 
y = b\Sin(t)\cos\phi \\
z =  b\Sin(t)\sin\phi
\end{align}

The azimutal angle where the line of sight is tangent to the shell is given by equation (\ref{eq:tanphi}) and (\ref{eq:tanalpha}), then:

\begin{align}
\sin\phi_t &= \frac{b}{a}\tan i\Cot(t) 
\end{align}
where:

\begin{align}
\Cot(t) = \left\lbrace \begin{array}{c}
\cot t ~\mathrm{if~ellipse} \\
-\coth t ~\mathrm{if~hyperbola}
\end{array}\right.
\end{align}

We could work in another frame where the origin is located at the conic's center: $(X,Y)$, where $X=x-X_0$ and $Y=y$.

In this frame,  we use the transformations (\ref{eq:Trans})  to obtain the apparent shape of the shell:

\begin{align}
X' = \frac{\Cos(t)}{a\cos i}\left(a^2\cos^2 i \pm b^2\sin^2 i\right)  \label{eq:conic-projected-x}\\
Y'= b\Sin(t)\left(1-\frac{b^2}{a^2}\tan^2 i\Cot(t)\right)^{1/2}
\label{eq:conic-projected-y}
\end{align}

We must show that the apparent shape of the shell in the observer's frame is still the same conic. To do that, we can write equations
(\ref{eq:conic-projected-x}) and (\ref{eq:conic-projected-y}) as follows:

\begin{align}
X' = a'\Cos t' \label{eq:conic-projected-x-2} \\
Y' = b'\Sin t' \label{eq:conic-projected-y-2}
\end{align}

Comparing, (\ref{eq:conic-projected-x}), (\ref{eq:conic-projected-x-2}), (\ref{eq:conic-projected-y}) and (\ref{eq:conic-projected-y}) we find that

\begin{align}
a' = \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2} \\
\Cos(t') = \frac{\Cos(t)}{a\cos i} \\
b' = b
\end{align}

Which proves that effectively the projection of a given conic is the same conic. With this knowledge, we can obtain the characteristic radii in
the observer's frame.

\begin{align}
D' = D\cos i \\
R'_0 = a' - (a-R_0)\cos i \\
R'_0 = \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2}  - (a-R_0)\cos i
\end{align}

With  $f(i;\theta_c)\equiv\left(1\pm\tan^2\theta_c\tan^2i\right)^{1/2}$ and introducing the
radius of curvature $R_c\equiv \frac{b^2}{a}$ we find that:

\begin{align}
\frac{R'_0}{D'}=\frac{R_0}{D}\left(1+A\cot^2\theta_c(f(i;\theta_c)-1) \right)
\end{align}

Where $A\equiv \frac{R_c}{R_0}$

The Radius of curvature in the observer's frame is given by $R'_c=\frac{b'^2}{a'}$. Then:

\begin{align}
\frac{R'_c}{R'_0} = \frac{A}{\cos^2 i f(i;\theta_c)\frac{q'}{q}}
\end{align}

Where $q=\frac{R_0}{D}$ and $q' = \frac{R'_0}{D'}$
\begin{figure*}
\includegraphics[width=0.8\linewidth]{conic2}
\label{fig:conic2}
\caption{Graphic example of the increasing of the projected radius of curvature with inclination}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=0.9\linewidth]{annotated}
\caption{Projected Radius of curvature vs projected $R_0$ for $\theta_c=60^\circ,45^\circ,28^\circ, 0^\circ,-15^\circ,-30^\circ$. Each colored points represent
the same shape at a different inclination. Being the dark blue the ones with the lower inclination and the red ones with the highest value. For closed shapes, there is
an assymptotic limit for the projected radius of curvature at high inclinations, while for open shapes there is an upper limit for inclination where a tangent line can be observed.}
\label{fig:Apqp}
\end{figure*}

>>>>

\section{Conic section approximation}
\label{sec:conic}

@@

%Instead of the excentricity, we utilize the parameter $\theta_c$ to characterize the different curves, where
%$\tan\theta_c = \frac{b}{a}$,  $b$ and $a$ are the typical parameters of conics. A positive value for $\theta_c$ indicates that the given curve is a closed one, i.e
%an ellipse, while a negative value indicates that is an hyperbola. %Insert figures if neccesary  
\begin{figure}
\setlength\tabcolsep{0pt}
\begin{tabular}{cc}
\includegraphics[height=0.44\linewidth, trim=30 0 30 0, clip]{ellipse_py_lab} &
\includegraphics[height=0.44\linewidth, trim=60 0 70 0, clip]{hyperbola_py}
\end{tabular}
\label{fig:conics}
\caption{Schematic representation of conic like bowshocks: A) Ellipse and B) Hyperbola The center of the conic is located at a distance $x_0$ from the origin at the $x$ axis. The shell position is 
described either by cartesian coordinates $(x,y)$ or polar coordinates $(R,\theta)$. In this section is more convenient to use cartesian coordinates, as their parametrization is quite easy.}
\end{figure}
\begin{figure}
\includegraphics[width=\linewidth]{conic1}
\caption{A whole set of conics with the same radius of curvature. The parameter $\theta_c$ is used as the equivalent of the excentricity. This angle is defined as positive for closed curves
and negative for the opened ones. An special case is $\theta_c=0$, which describes a parabola.}
\label{fig:conics-family}
\end{figure}

Due to the similarities between the parametrization of the ellipse and the hyperbola, we can do the following parametrization:
\begin{align}
x = a\Cos(t)-x_0 \label{eq:par-x}\\ 
y = b\Sin(t) \label{eq:par-y}
\end{align}
where:
\begin{align}
\Cos(t) = \left\lbrace \begin{array}{c}
\cos t ~\mathrm{if~ellipse} \\
\cosh t ~\mathrm{if~hyperbola}
\end{array}\right.\\
\Sin(t) = \left\lbrace \begin{array}{c}
\sin t ~\mathrm{if~ellipse} \\
\sinh t ~\mathrm{if~hyperbola}
\end{array}\right. \\
-\pi < t < \pi \\
R_0 = a - x_0 \\
a<0 ~ \mathrm{for~hyperbola}
\end{align}
We have set a whole family of curves with a singular parameter denoted as $\tan\theta_c \equiv \frac{b}{a}$. Positive values of this parameter describe closed curves
(i.e. ellipses) and the negative ones describe open curves (i.e. hyperbolas). Particular cases are $\theta_c =\frac{\pi}{4}$, which describes a circle and $\theta_c=0$, which describes
a parabola.

\subsection{Plane-of-sky projection of conic sections} 
\begin{figure*}
\includegraphics[width=0.9\linewidth]{annotated}
\caption{Projected Radius of curvature vs projected $R_0$ for $\theta_c=60^\circ,45^\circ,28^\circ, 0^\circ,-15^\circ,-30^\circ$. Each colored points represent
the same shape at a different inclination. Being the dark blue the ones with the lower inclination and the red ones with the highest value. For closed shapes, there is
an assymptotic limit for the projected radius of curvature at high inclinations, while for open shapes there is an upper limit for inclination where a tangent line can be observed.}
\label{fig:Apqp}
\end{figure*}

Once the parametrization is done, we can find the apparent shape of
the shell in the observer's frame, following the procedure explained
in section \ref{sec:projection}.  First of all, the intrinsic 3D shape of the shell is given by:
\begin{align}
x = a\Cos(t)-x_0 \\ 
y = b\Sin(t)\cos\phi \\
z =  b\Sin(t)\sin\phi
\end{align}
The azimutal angle where the line of sight is tangent to the shell is given by equation (\ref{eq:tanphi}) and (\ref{eq:tanalpha}), then:
\begin{align}
\sin\phi_t &= \frac{b}{a}\tan i\Cot(t) 
\end{align}
where:
\begin{align}
\Cot(t) = \left\lbrace \begin{array}{c}
\cot t ~\mathrm{if~ellipse} \\
-\coth t ~\mathrm{if~hyperbola}
\end{array}\right.
\end{align}

We could work in another frame where the origin is located at the conic's center: $(X,Y)$, where $X=x-X_0$ and $Y=y$.
In this frame,  we use the transformations (\ref{eq:Trans})  to obtain the apparent shape of the shell:
\begin{align}
X' = \frac{\Cos(t)}{a\cos i}\left(a^2\cos^2 i \pm b^2\sin^2 i\right)  \label{eq:conic-projected-x}\\
Y'= b\Sin(t)\left(1-\frac{b^2}{a^2}\tan^2 i\Cot(t)\right)^{1/2}
\label{eq:conic-projected-y}
\end{align}

We must show that the apparent shape of the shell in the observer's frame is still the same conic. To do that, we can write equations
(\ref{eq:conic-projected-x}) and (\ref{eq:conic-projected-y}) as follows:
\begin{align}
X' = a'\Cos t' \label{eq:conic-projected-x-2} \\
Y' = b'\Sin t' \label{eq:conic-projected-y-2} . 
\end{align}

Comparing, (\ref{eq:conic-projected-x}), (\ref{eq:conic-projected-x-2}), (\ref{eq:conic-projected-y}) and (\ref{eq:conic-projected-y}) we find that
\begin{align}
a' = \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2} \\
\Cos(t') = \frac{\Cos(t)}{a\cos i} \\
b' = b
\end{align} 
Which proves that effectively the projection of a given conic is the same conic. With this knowledge, we can obtain the characteristic radii in
the observer's frame.
\begin{align}
D' = D\cos i \\
R'_0 = a' - (a-R_0)\cos i \\
R'_0 = \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2}  - (a-R_0)\cos i
\end{align}

With  $f(i;\theta_c)\equiv\left(1\pm\tan^2\theta_c\tan^2i\right)^{1/2}$ and introducing the
radius of curvature $R_c\equiv \frac{b^2}{a}$ we find that:
\begin{align}
\frac{R'_0}{D'}=\frac{R_0}{D}\left(1+A\cot^2\theta_c(f(i;\theta_c)-1) \right)
\end{align}
Where $A\equiv \frac{R_c}{R_0}$

The Radius of curvature in the observer's frame is given by $R'_c=\frac{b'^2}{a'}$. Then:
\begin{align}
\frac{R'_c}{R'_0} = \frac{A}{\cos^2 i f(i;\theta_c)\frac{q'}{q}}
\end{align}
where $q=\frac{R_0}{D}$ and $q' = \frac{R'_0}{D'}$. 
% \begin{figure*}
% \includegraphics[width=0.8\linewidth]{conic2}
% \label{fig:conic2}
% \caption{Graphic example of the increasing of the projected radius of curvature with inclination}
% \end{figure*}

@@

Finally, we need to obtain $R_{90}$ using equations (\ref{eq:par-x}) and (\ref{eq:par-y}):
\begin{align}
x_{90} = a - R_0 = a\Cos(t_{90}) \\
y_{90} = R_{90} = b\Sin(t_{90})
\end{align}
With this we find that:
\begin{align}
R_{90} = b\left(1 \mp \frac{(R_0-a)^2}{a^2}\right)
\end{align}
